Towards Explainability of Machine Learning Models in Insurance Pricing

### 2019-10-24 manuscript/manuscript.Rmd
use elsevier template

---
title: |
 Towards Explainability of Machine Learning Models in Insurance Pricing
type: WIP DRAFT 
author:
  - name: Kevin Kuo
    affil: a
    email: kevin@kasa.ai
  - name: Daniel Lupton
    affil: b
    email: dlupton@taylorandmulder.com
affiliation:
  - num: a
    address: |
      Kasa AI,
      3040 78th Ave SE #1271,
      Mercer Island, WA 98040, USA
  - num: b
    address: |
      Taylor & Mulder,
      10508 Rivers Bend Lane,
      Potomac, MD 20854, USA
bibliography: explain-pricing.bib
appendix: appendix.tex
citation_package: bibtex
abstract: |
  Abstract tbd
keywords: |
  actuarial science; general insurance
header-includes: |
  \usepackage{hyperref}
  \usepackage[utf8]{inputenc}
  \def\tightlist{}
output: 
  bookdown::pdf_book:
    base_format: rticles::tf_article
    number_sections: true
---

@@

Paralelling the development of GLM was the development of machine learning algorithms throughout the middle part of the 20th century. Detailed histories of machine learning may be found in sources such as @Nilsson_2009 and @wang_raj_2017. Consistent with GLMs, machine learning was relatively unpopular in actuarial science until the last ten years as computing power has become cheaper and more easily available and as machine learning software packages have obviated the need for developing analyses from scratch each time an analysis is performed. Due to the breadth of machine learning as a field, it is difficult to identify the first time it entered the CAS syllabus; however, cluster analysis (in the form of k-means) seems to have been first included in 2011 with @robertson_2009. More recently, the MAS-I and MAS-II exams introduced in 2018 have included machine learning explicitly.

@@

```{r tree-plot1, echo = FALSE, fig.cap="A simple decision tree for loss cost prediction.", out.width="70%", fig.align='center'}
knitr::include_graphics("figures/tree-plot1.png")
```

@@

```{r tree-plot2, echo = FALSE, fig.cap="A more complex decision tree for loss cost prediction.", out.width="90%", fig.align='center'}
knitr::include_graphics("figures/tree-plot2.png")
```

@@

```{r fi, out.width='60%', fig.align = "center", fig.cap="Permutation feature importances for ML model.", echo = FALSE}
knitr::include_graphics("figures/fi-plot.png")
```

@@

@@

```{r pdp, out.width='60%', fig.align = "center", fig.cap="Partial dependence plot for ML model.", echo = FALSE}
knitr::include_graphics("figures/pdp-plot.png")
```

@@

```{r breakdown, out.width='80%', fig.align = "center", fig.cap="Variable contribution plot for ML model.", echo = FALSE}
knitr::include_graphics("figures/breakdown-plot.png")
```

@@

>>>>

---
title: Towards Explainability of Machine Learning Models in Insurance Pricing
author:
  - name: Kevin Kuo
    email: kevin@kasa.ai
    affiliation: a
    footnote: Corresponding Author
  - name: Daniel Lupton
    email: dlupton@taylorandmulder.com
    affiliation: b
address:
  - code: a
    address: Kasa AI, 3040 78th Ave SE \#1271, Mercer Island, WA 98040, USA
  - code: b
    address: Taylor & Mulder, 10508 Rivers Bend Lane, Potomac, MD 20854, USA
abstract: |
  Machine learning methods have garnered increasing interest among actuaries in recent years. However, their adoption by practitioners has been limited, partly due to the lack of transparency of these methods, as compared to generalized linear models. In this paper, we discuss the need for model interpretability in property & casualty insurance ratemaking, propose a framework for explaining models, and present a case study to illustrate the framework.

journal: "[journal]"
date: "`r Sys.Date()`"
bibliography: explain-pricing.bib
#linenumbers: true
numbersections: true
# csl: elsevier-harvard.csl
output: 
  bookdown::pdf_book:
    base_format: rticles::elsevier_article
layout: preprint, 3p, twocolumn, letterpaper, 12pt
---

@@

Paralelling the development of GLM was the development of machine learning algorithms throughout the middle part of the 20th century. Detailed histories of machine learning may be found in sources such as @nilsson_2009 and @wang_raj_2017. Consistent with GLMs, machine learning was relatively unpopular in actuarial science until the last ten years as computing power has become cheaper and more easily available and as machine learning software packages have obviated the need for developing analyses from scratch each time an analysis is performed. Due to the breadth of machine learning as a field, it is difficult to identify the first time it entered the CAS syllabus; however, cluster analysis (in the form of k-means) seems to have been first included in 2011 with @robertson_2009. More recently, the MAS-I and MAS-II exams introduced in 2018 have included machine learning explicitly.

@@

```{r tree-plot1, echo = FALSE, fig.cap="A simple decision tree for loss cost prediction.", out.width="100%", fig.align='center'}
knitr::include_graphics("figures/tree-plot1.png")
```

@@

```{r tree-plot2, echo = FALSE, fig.cap="A more complex decision tree for loss cost prediction.", out.width="60%", fig.align='center',  fig.env='figure*'}
knitr::include_graphics("figures/tree-plot2.png")
```

@@

```{r fi, out.width='100%', fig.align = "center", fig.cap="Permutation feature importances for ML model.", echo = FALSE}
knitr::include_graphics("figures/fi-plot.png")
```

@@

```{r pdp, out.width='70%', fig.align = "center", fig.cap="Partial dependence plot for ML model.", echo = FALSE,  fig.env='figure*', fig.pos='h'}
knitr::include_graphics("figures/pdp-plot.png")
```

```{r breakdown, out.width='80%', fig.align = "center", fig.cap="Variable contribution plot for ML model.", echo = FALSE, fig.env = 'figure*', fig.pos = 'h'}
knitr::include_graphics("figures/breakdown-plot.png")
```

@@

@@

@@

References {#references .unnumbered}
==========
<div id="refs"></div>

\appendix

Appendix {.unnumbered}
=============

# ML Model Development

In this appendix, we describe the ML model and the data used to train it. Note that, for our paper, the ultimate goal of the modeling procedure is to develop something that can produce predictions. As a result, we do not follow standard practices for supervised learning model development. However, for the sake of completeness and reproducibility, we include the details of the process here.

## Data

We use data from the AUTOSEG ("Automobile Statistics System") of Brazil's Superintendence of Private Insurance (SUSEP). The organization maintains policy-characteristics-level data, including claim counts and amounts, for all insured automobiles in Brazil. The data contains variables from policyholder characteristics to losses by peril.

## Model

Model
